\textbf{ Accurate spectrophotometric estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-27 paper/sp.tex
some intro text

The promise of mapping the Milky Way disk with infrared spectroscopy
and red-giant stars.

The information \gaia\ brings to this.

>>>>

If we want to make precise kinematic and element-abundance maps
of the Milky Way disk out to large heliocentric radii,
and especially through the extinction to the other side of the Galactic Center, we will
need to make use of luminous red giants, and we will need to observe in
the infrared.
These arguments motivate the \apogee\ (CITE), \apogee-2 (CITE), and \sdssv\ (CITE)
projects, which take spectra in the infrared, and deliver detailed abundances
along the entire red-giant branch up to the tip.
The maps made with these data will reveal the dynamics of the Milky Way and its disk,
and show us how and where stars form, and how they migrate around (and out of) the
disk with time.

As we operate these spectroscopic surveys,
the \gaia\ Mission has been revolutionizing our view of the disk.
It delivers good proper motions everywhere,
and extremely valuable parallax information
locally.
However, it does not deliver precise geometric parallaxes over a large fraction of
the disk, and certainly not in the dusty and crowded regions.
For these reasons, the value of \gaia\ in these infrared sky projects is not
to deliver distance information directly, but rather to calibrate stellar models,
which then deliver distance information through relationships between stellar
luminosities and spectral characteristics.

In this contribution we develop data-driven techniques to predict
stellar luminosities and distances with spectroscopy and photometry...
